---
title: "Ontario Greenhouse Gas Resampling Efforts"
author: "Brandon Edwards"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(cmdstanr)
library(ggplot2)
library(ggpubr)
library(grid)
library(bayesplot)
theme_set(theme_pubclean())
```
# Executive Summary

The purpose of this report is to determine whether the 16 restored wetland sites in southern Ontario can be sampled 3 times in 2025, rather than the usual 5 times as done in 2023 and 2024. Here, I have run a series of Bayesian models to determine this. I first run a series of models on the full data set to assess modelled effects under full data. Then, I run another series of models under a reduced data scenario to assess modelled effects under reduced data. Using Bayesian posterior analysis, I then show that there is very little difference in modelled effects between the full data set and the reduced data set. One exception is models of N2O at site ON12, where the reduced data set appears to have lost sufficient enough information such that the statistical effect at that site goes to 0. With these results, I am comfortable recommending that sampling be reduced from 5 visits per year to 3 visits per year, with the caveat that some sites may experience *some* information loss. However, careful modelling of the data later on can help recover some of the lost information with the reduced sampling, by borrowing information from the more highly sampled years using a Bayesian modelling framework.

# Analysis

```{r data_cleaning}
data <- read.csv("data/raw/data_full.csv")

# Remove extra empty columns
data <- data[, 1:8]

#' Add visit number as a proxy for the date. We'll see later on if we want
#' the actual date or not. But, 1 - 5 will correspond with early to late
data$Visit <- rep(seq(1,5), times = nrow(data) / 5)
data$Visit <- factor(data$Visit, levels = c("1", "2", "3", "4", "5"))

# Must also make sure the values are numeric again
data$pCO2.corrected <- as.numeric(data$pCO2.corrected)
data$pCH4.insitu <- as.numeric(data$pCH4.insitu)
data$pN2O.insitu <- as.numeric(data$pN2O.insitu)

# And make the year a factor
data$Year <- factor(data$Year, levels = c("2023", "2024"))
```

## Visualize Data

#### Greenhouse Gas by Visit x Year Combination

Before running any analyses, let's see how the different greenhouse gases vary through the year (via the visit number), and across the years. We will plot the site visits and year simultaneously as boxplots, such that we have visit number on the x axis, the log greenhouse gas amount on the y axis, and boxplots coloured by year. This will allow us to see if there is any particular interactions we should be thinking about when considering future sample size amounts.

Let's first look at CO2.

```{r plot_co2_visityear, echo=FALSE}

ggplot(data = data, aes(x = Visit, y = log(pCO2.corrected))) +
  geom_boxplot(aes(fill = Year)) +
  #facet_wrap(~Site) +
  NULL

```


Now for CH4

```{r plot_ch4_visityear, echo=FALSE}

ggplot(data = data, aes(x = Visit, y = log(pCH4.insitu))) +
  geom_boxplot(aes(fill = Year)) +
  #facet_wrap(~Site) +
  NULL

```


Now for N2O

```{r plot_n2o_visityear, echo=FALSE}

ggplot(data = data, aes(x = Visit, y = log(pN2O.insitu))) +
  geom_boxplot(aes(fill = Year)) +
  #facet_wrap(~Site) +
  NULL
```

#### Greenhouse Gas by Site x Year Combination

We can also see how the sites vary by year with greenhouse gas amounts. Let's first start with CO2:

```{r plot_co2_siteyear, echo=FALSE}

ggplot(data = data, aes(x = Year, y = log(pCO2.corrected))) +
  geom_boxplot(aes(fill = Year)) +
  facet_wrap(~Site)


```

And now for CH4:

```{r plot_ch4_siteyear, echo=FALSE}

ggplot(data = data, aes(x = Year, y = log(pCH4.insitu))) +
  geom_boxplot(aes(fill = Year)) +
  facet_wrap(~Site)

```

And finally for N2O:

```{r plot_n2o_siteyear, echo=FALSE}

ggplot(data = data, aes(x = Year, y = log(pN2O.insitu))) +
  geom_boxplot(aes(fill = Year)) +
  facet_wrap(~Site)

```

#### Greenhouse Gases by Site x Visit x Year

Out of curiosity, we could see how the greenhouse gas amounts change through the year, between years, at each of the sites. Since there is only 1 sample per year x visit x site combination, we won't be able to model at that level, but it would be worth looking at.

We will first make a plot of CO2 versus visit, for each site, coloured by year.

```{r, echo=FALSE}

ggplot(data = data, aes(x = Visit, y = log(pCO2.corrected), group = Year, color = Year)) +
  geom_line() +
  facet_wrap(~Site)


```


Now for CH4

```{r, echo=FALSE}

ch4_plot <- ggplot(data = data, aes(x = Visit, y = log(pCH4.insitu), group = Year, color = Year)) +
  geom_line() +
  facet_wrap(~Site)

print(ch4_plot)

```


Now for N2O

```{r, echo=FALSE}

n2o_plot <- ggplot(data = data, aes(x = Visit, y = log(pN2O.insitu), group = Year, color = Year)) +
  geom_line() +
  facet_wrap(~Site)

print(n2o_plot)

```


## Modelling Approach

Based on the plots above, I am seeing that there are two different questions we could ask of the data:

1) How do greenhouse gas concentrations change over a given year, and does this change from year to year? This is based on the Greenhouse Gas by Visit x Year Combination plots above.
2) How do greenhouse gas concentrations change at a given site, and does this change from year to year? This is based on the Greenhouse Gas by Site x Year Combination plots above. 

Given either of these questions, we want to know if we can answer these questions using only 3 samples from each site in 2025, rather than 5 samples from each site. To accomplish this, I will take on the following modelling approach:

1) Run each combination above as a Bayesian mixed effects model. That is, for Question 1, I will treat year as a random effect, and visit as a fixed effect; for Question 2, I will treat year as a random effect, and site as a fixed effect. All models will include an interaction term. I will therefore generate 6 models.
2) Analyze the posterior distribution of the parameters for each model. To do this, I will plot the median value with the 50% and 90% credible intervals.
3) Then, reduce the dataset to three visits per site. Here, I will choose to retain Visit 1, Visit 3, and Visit 5, which would correspond to sampling in Spring, Summer, and Fall at each site. I will then run the same models as 1) but with the reduced data set.
4) Using Bayesian posterior analysis, we can analyse the *distribution* of the difference in parameter estimates between two (or more) models. Here, I will analyse the distribution of the difference in parameter estimates for the set of models for Question 1 and the set of models for Question 2. If we detect little to no changes in posterior distributions of parameter estimates (i.e., the distribution of the difference in parameter estimates is approximately 0), then we can conclude that we are **not** missing much information when we reduce our sampling from 5 per site to 3 visits per site.

All models will be done in Stan, using 2000 warmup iterations and 1000 sampling iterations on 4 chains, for a total of 4000 posterior draws. Model code can be found here: https://github.com/DUC-IWWR/ontario-ghg-resampling

Let's now dive into these models.

### Full Data

#### How do greenhouse gas concentrations change over a given year, and does this change from year to year?

```{r model-setup, echo=FALSE}

data_red <- data[-which(data$Still.missing.... == "*"), ]
data_red$Year_Visit <- paste0(data_red$Year, "-", data_red$Visit)
data_red$Year_Site <- paste0(data_red$Year, "-", data_red$Site)

model <- cmdstan_model("models/model.stan")

mod1_plot_labels <- c("Mean", "Year_Mean", "2023", "2024", "Visit 1", "Visit 2",
                      "Visit 3", "Visit 4", "Visit 5", "Interaction", "2023_1",
                      "2023_2", "2023_3", "2023_4", "2023_5", "2024_1",
                      "2024_2", "2024_3", "2024_4", "2024_5")
mod2_plot_labels <- c("Mean", "Year_Mean", "2023", "2024", unique(data_red$Site),
                      "Interaction", unique(data_red$Year_Site))
```

```{r co2_mod1, echo = FALSE}
co2_data <- list(N = nrow(data_red),
                 y = log(data_red$pCO2.corrected),
                 n_factor1 = length(unique(data_red$Year)),
                 factor1 = as.numeric(as.factor(data_red$Year)),
                 n_factor2 = length(unique(data_red$Visit)),
                 factor2 = data_red$Visit,
                 n_interaction = length(unique(data_red$Year_Visit)),
                 interaction = as.numeric(as.factor(data_red$Year_Visit)))

co2_mod1 <- model$sample(
  data = co2_data,
  iter_warmup = 2000,
  iter_sampling = 1000,
  chains = 4,
  parallel_chains = 4,
  refresh = 10,
  show_messages = FALSE,
  show_exceptions = FALSE
)

co2_mod1_plot <- bayesplot::mcmc_intervals(co2_mod1$draws(c("mu", "alpha_hyper", "alpha", "beta", "gamma_hyper", "gamma"))) +
  scale_y_discrete(labels = mod1_plot_labels)

```

```{r ch4_mod1, echo = FALSE}
ch4_data <- list(N = nrow(data_red),
                 y = log(data_red$pCH4.insitu),
                 n_factor1 = length(unique(data_red$Year)),
                 factor1 = as.numeric(as.factor(data_red$Year)),
                 n_factor2 = length(unique(data_red$Visit)),
                 factor2 = data_red$Visit,
                 n_interaction = length(unique(data_red$Year_Visit)),
                 interaction = as.numeric(as.factor(data_red$Year_Visit)))

ch4_mod1 <- model$sample(
  data = ch4_data,
  iter_warmup = 2000,
  iter_sampling = 1000,
  chains = 4,
  parallel_chains = 4,
  refresh = 10,
  show_messages = FALSE,
  show_exceptions = FALSE
)

ch4_mod1_plot <- bayesplot::mcmc_intervals(ch4_mod1$draws(c("mu", "alpha_hyper", "alpha", "beta", "gamma_hyper", "gamma"))) +
  scale_y_discrete(labels = mod1_plot_labels) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

```{r n2o_mod1, echo = FALSE}
n2o_data <- list(N = nrow(data_red),
                 y = log(data_red$pN2O.insitu),
                 n_factor1 = length(unique(data_red$Year)),
                 factor1 = as.numeric(as.factor(data_red$Year)),
                 n_factor2 = length(unique(data_red$Visit)),
                 factor2 = data_red$Visit,
                 n_interaction = length(unique(data_red$Year_Visit)),
                 interaction = as.numeric(as.factor(data_red$Year_Visit)))

n2o_mod1 <- model$sample(
  data = n2o_data,
  iter_warmup = 2000,
  iter_sampling = 1000,
  chains = 4,
  parallel_chains = 4,
  refresh = 10,
  show_messages = FALSE,
  show_exceptions = FALSE
)

n2o_mod1_plot <- bayesplot::mcmc_intervals(n2o_mod1$draws(c("mu", "alpha_hyper", "alpha", "beta", "gamma_hyper", "gamma"))) +
    scale_y_discrete(labels = mod1_plot_labels) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

Here, I will plot the posterior distributions of all relevant effects for the three greenhouses gases. We will have a distribution for the overall mean concentration, a yearly mean plus two year effects, a visit effect, and an interaction mean plus individual interaction effects. The circles are the median estimates for each effect; the bold line is the 50% credible interval of the posterior distribution, and the thin line is the 90% credible interval of the posterior distribution.

```{r mod1_plots, echo=FALSE}
ggarrange(co2_mod1_plot, ch4_mod1_plot + rremove("ylab"), n2o_mod1_plot + rremove("ylab"),
          labels = c("CO2", "CH4", "N2O"),
          nrow = 1,
          widths = c(0.4,0.3,0.3))
```

#### How do greenhouse gas concentrations change at a given site, and does this change from year to year?



```{r co2_mod2, echo = FALSE}
co2_data2 <- list(N = nrow(data_red),
                 y = log(data_red$pCO2.corrected),
                 n_factor1 = length(unique(data_red$Year)),
                 factor1 = as.numeric(as.factor(data_red$Year)),
                 n_factor2 = length(unique(data_red$Site)),
                 factor2 = as.numeric(as.factor(data_red$Site)),
                 n_interaction = length(unique(data_red$Year_Site)),
                 interaction = as.numeric(as.factor(data_red$Year_Site)))

co2_mod2 <- model$sample(
  data = co2_data2,
  iter_warmup = 2000,
  iter_sampling = 1000,
  chains = 4,
  parallel_chains = 4,
  refresh = 10,
  show_messages = FALSE,
  show_exceptions = FALSE
)

co2_mod2_plot <- bayesplot::mcmc_intervals(co2_mod2$draws(c("mu", "alpha_hyper", "alpha", "beta", "gamma_hyper", "gamma"))) +
  scale_y_discrete(labels = mod2_plot_labels)

```

```{r ch4_mod2, echo = FALSE}
ch4_data2 <- list(N = nrow(data_red),
                 y = log(data_red$pCH4.insitu),
                 n_factor1 = length(unique(data_red$Year)),
                 factor1 = as.numeric(as.factor(data_red$Year)),
                 n_factor2 = length(unique(data_red$Site)),
                 factor2 = as.numeric(as.factor(data_red$Site)),
                 n_interaction = length(unique(data_red$Year_Site)),
                 interaction = as.numeric(as.factor(data_red$Year_Site)))

ch4_mod2 <- model$sample(
  data = ch4_data2,
  iter_warmup = 2000,
  iter_sampling = 1000,
  chains = 4,
  parallel_chains = 4,
  refresh = 10,
  show_messages = FALSE,
  show_exceptions = FALSE
)

ch4_mod2_plot <- bayesplot::mcmc_intervals(ch4_mod2$draws(c("mu", "alpha_hyper", "alpha", "beta", "gamma_hyper", "gamma"))) +
  scale_y_discrete(labels = mod2_plot_labels) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

```{r n2o_mod2, echo = FALSE}
n2o_data2 <- list(N = nrow(data_red),
                 y = log(data_red$pN2O.insitu),
                 n_factor1 = length(unique(data_red$Year)),
                 factor1 = as.numeric(as.factor(data_red$Year)),
                 n_factor2 = length(unique(data_red$Site)),
                 factor2 = as.numeric(as.factor(data_red$Site)),
                 n_interaction = length(unique(data_red$Year_Site)),
                 interaction = as.numeric(as.factor(data_red$Year_Site)))

n2o_mod2 <- model$sample(
  data = n2o_data2,
  iter_warmup = 2000,
  iter_sampling = 1000,
  chains = 4,
  parallel_chains = 4,
  refresh = 10,
  show_messages = FALSE,
  show_exceptions = FALSE
)

n2o_mod2_plot <- bayesplot::mcmc_intervals(n2o_mod2$draws(c("mu", "alpha_hyper", "alpha", "beta", "gamma_hyper", "gamma"))) +
  scale_y_discrete(labels = mod2_plot_labels) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```
For this model, we will now have a distribution for the overall mean concentration, a yearly mean plus two year effects, site effects, and an interaction mean plus individual interaction effects. The circles are the median estimates for each effect; the bold line is the 50% credible interval of the posterior distribution, and the thin line is the 90% credible interval of the posterior distribution.

```{r mod2_plots, echo=FALSE}
ggarrange(co2_mod2_plot, ch4_mod2_plot + rremove("ylab"), n2o_mod2_plot + rremove("ylab"),
          labels = c("CO2", "CH4", "N2O"),
          nrow = 1,
          widths = c(0.4,0.3,0.3))
```
### Reduced Data

Now that we see what the posterior distributions look like here, we can actually try a scenario where we get rid of Visit 2 and Visit 4 (which would correspond to planned visits in Spring/Summer/Fall), and assess how much less precise our estimates get. Let's try this now. The following plots will be of the same models as above, but with Visit 2 and Visit 4 removed from the entire study.

```{r data-reduction, echo=FALSE}

data_red <- data_red[which(data_red$Visit %in% c("1", "3", "5")), ]
data_red[which(data_red$Visit == 3), "Visit"] <- 2
data_red[which(data_red$Visit == 5), "Visit"] <- 3
data_red$Visit <- as.numeric(as.factor(data_red$Visit))

mod1_3visit_plot_labels <- c("Mean", "Year_Mean", "2023", "2024", "Visit 1", 
                      "Visit 3", "Visit 5", "Interaction", "2023_1",
                       "2023_3", "2023_5", "2024_1",
                       "2024_3", "2024_5")
mod2_3visit_plot_labels <- c("Mean", "Year_Mean", "2023", "2024", unique(data_red$Site),
                      "Interaction", unique(data_red$Year_Site))

#model <- cmdstan_model("models/model.stan")
```

```{r co2_mod1_3visit, echo = FALSE}
co2_data_3visit <- list(N = nrow(data_red),
                 y = log(data_red$pCO2.corrected),
                 n_factor1 = length(unique(data_red$Year)),
                 factor1 = as.numeric(as.factor(data_red$Year)),
                 n_factor2 = length(unique(data_red$Visit)),
                 factor2 = data_red$Visit,
                 n_interaction = length(unique(data_red$Year_Visit)),
                 interaction = as.numeric(as.factor(data_red$Year_Visit)))

co2_mod1_3visit <- model$sample(
  data = co2_data_3visit,
  iter_warmup = 2000,
  iter_sampling = 1000,
  chains = 4,
  parallel_chains = 4,
  refresh = 10,
  show_messages = FALSE,
  show_exceptions = FALSE
)

co2_mod1_3visit_plot <- bayesplot::mcmc_intervals(co2_mod1_3visit$draws(c("mu", "alpha_hyper", "alpha", "beta", "gamma_hyper", "gamma"))) +
  scale_y_discrete(labels = mod1_3visit_plot_labels)

```

```{r ch4_mod1_3visit, echo = FALSE}
ch4_data <- list(N = nrow(data_red),
                 y = log(data_red$pCH4.insitu),
                 n_factor1 = length(unique(data_red$Year)),
                 factor1 = as.numeric(as.factor(data_red$Year)),
                 n_factor2 = length(unique(data_red$Visit)),
                 factor2 = data_red$Visit,
                 n_interaction = length(unique(data_red$Year_Visit)),
                 interaction = as.numeric(as.factor(data_red$Year_Visit)))

ch4_mod1_3visit <- model$sample(
  data = ch4_data,
  iter_warmup = 2000,
  iter_sampling = 1000,
  chains = 4,
  parallel_chains = 4,
  refresh = 10,
  show_messages = FALSE,
  show_exceptions = FALSE
)

ch4_mod1_3visit_plot <- bayesplot::mcmc_intervals(ch4_mod1_3visit$draws(c("mu", "alpha_hyper", "alpha", "beta", "gamma_hyper", "gamma"))) +
  scale_y_discrete(labels = mod1_3visit_plot_labels) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

```{r n2o_mod1_3visit, echo = FALSE}
n2o_data <- list(N = nrow(data_red),
                 y = log(data_red$pN2O.insitu),
                 n_factor1 = length(unique(data_red$Year)),
                 factor1 = as.numeric(as.factor(data_red$Year)),
                 n_factor2 = length(unique(data_red$Visit)),
                 factor2 = data_red$Visit,
                 n_interaction = length(unique(data_red$Year_Visit)),
                 interaction = as.numeric(as.factor(data_red$Year_Visit)))

n2o_mod1_3visit <- model$sample(
  data = n2o_data,
  iter_warmup = 2000,
  iter_sampling = 1000,
  chains = 4,
  parallel_chains = 4,
  refresh = 10,
  show_messages = FALSE,
  show_exceptions = FALSE
)

n2o_mod1_3visit_plot <- bayesplot::mcmc_intervals(n2o_mod1_3visit$draws(c("mu", "alpha_hyper", "alpha", "beta", "gamma_hyper", "gamma"))) +
  scale_y_discrete(labels = mod1_3visit_plot_labels) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

```{r mod1_plots_3visit, echo=FALSE}
ggarrange(co2_mod1_3visit_plot, ch4_mod1_3visit_plot + rremove("ylab"), n2o_mod1_3visit_plot + rremove("ylab"),
          labels = c("CO2", "CH4", "N2O"),
          nrow = 1,
          widths = c(0.4,0.3,0.3))
```





```{r co2_mod2_3visit, echo = FALSE}
co2_data2 <- list(N = nrow(data_red),
                 y = log(data_red$pCO2.corrected),
                 n_factor1 = length(unique(data_red$Year)),
                 factor1 = as.numeric(as.factor(data_red$Year)),
                 n_factor2 = length(unique(data_red$Site)),
                 factor2 = as.numeric(as.factor(data_red$Site)),
                 n_interaction = length(unique(data_red$Year_Site)),
                 interaction = as.numeric(as.factor(data_red$Year_Site)))

co2_mod2_3visit <- model$sample(
  data = co2_data2,
  iter_warmup = 2000,
  iter_sampling = 1000,
  chains = 4,
  parallel_chains = 4,
  refresh = 10,
  show_messages = FALSE,
  show_exceptions = FALSE
)

co2_mod2_3visit_plot <- bayesplot::mcmc_intervals(co2_mod2_3visit$draws(c("mu", "alpha_hyper", "alpha", "beta", "gamma_hyper", "gamma"))) +
  scale_y_discrete(labels = mod2_3visit_plot_labels) 

```

```{r ch4_mod2_3visit, echo = FALSE}
ch4_data2 <- list(N = nrow(data_red),
                 y = log(data_red$pCH4.insitu),
                 n_factor1 = length(unique(data_red$Year)),
                 factor1 = as.numeric(as.factor(data_red$Year)),
                 n_factor2 = length(unique(data_red$Site)),
                 factor2 = as.numeric(as.factor(data_red$Site)),
                 n_interaction = length(unique(data_red$Year_Site)),
                 interaction = as.numeric(as.factor(data_red$Year_Site)))

ch4_mod2_3visit <- model$sample(
  data = ch4_data2,
  iter_warmup = 2000,
  iter_sampling = 1000,
  chains = 4,
  parallel_chains = 4,
  refresh = 10,
  show_messages = FALSE,
  show_exceptions = FALSE
)

ch4_mod2_3visit_plot <- bayesplot::mcmc_intervals(ch4_mod2_3visit$draws(c("mu", "alpha_hyper", "alpha", "beta", "gamma_hyper", "gamma"))) +
  scale_y_discrete(labels = mod2_3visit_plot_labels) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

```{r n2o_mod2_3visit, echo = FALSE}
n2o_data2 <- list(N = nrow(data_red),
                 y = log(data_red$pN2O.insitu),
                 n_factor1 = length(unique(data_red$Year)),
                 factor1 = as.numeric(as.factor(data_red$Year)),
                 n_factor2 = length(unique(data_red$Site)),
                 factor2 = as.numeric(as.factor(data_red$Site)),
                 n_interaction = length(unique(data_red$Year_Site)),
                 interaction = as.numeric(as.factor(data_red$Year_Site)))

n2o_mod2_3visit <- model$sample(
  data = n2o_data2,
  iter_warmup = 2000,
  iter_sampling = 1000,
  chains = 4,
  parallel_chains = 4,
  refresh = 10,
  show_messages = FALSE,
  show_exceptions = FALSE
)

n2o_mod2_3visit_plot <- bayesplot::mcmc_intervals(n2o_mod2_3visit$draws(c("mu", "alpha_hyper", "alpha", "beta", "gamma_hyper", "gamma"))) +
  scale_y_discrete(labels = mod2_3visit_plot_labels) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

```{r mod2_plots_3visit, echo=FALSE}
ggarrange(co2_mod2_3visit_plot, ch4_mod2_3visit_plot + rremove("ylab"), n2o_mod2_3visit_plot + rremove("ylab"),
          labels = c("CO2", "CH4", "N2O"),
          nrow = 1,
          widths = c(0.4,0.3,0.3))
```
### Comparing the Full and Reduced Dataset using Bayesian Posterior Analysis

Visually, there does not appear to be large differences between the distributions of the reduced models compared to the distributions of the full models. We can see this especially in the model that considers site effects (i.e., question 2), in that the range of site-level effects and interactions are still nicely picked up with a reduced data set compared to the full data set. This visual inspect would imply that the amount of information contained in a 3-visit dataset is fairly similar to the amount of information contained in a 5-visit dataset.

We can, however, confirm this by taking the differences in the distribution of parameter estimates, such that we get a distribution of parameter differences. What we would expect to see is all parameter differences being approximately 0, meaning that we are not losing important information by sampling 3 times instead of 5 times. To contrast, if we saw differences different from 0, that would imply that we are effectively modelling different populations, and that we are missing something from sampling 3 times. 

Let's check for the first modelling question:

```{r mod1_difference, echo = FALSE}
co2_mod1_draws <- co2_mod1$draws(variables = c("mu", "alpha_hyper", "alpha", "beta[1]", "beta[3]", "beta[5]", "gamma_hyper",
                                               "gamma[1]", "gamma[3]", "gamma[5]", "gamma[6]", "gamma[8]", "gamma[10]"))
co2_mod1_3visit_draws <- co2_mod1_3visit$draws(variables = c("mu", "alpha_hyper", "alpha", "beta", "gamma_hyper", "gamma"))
co2_mod1_difference_plot <- bayesplot::mcmc_intervals(co2_mod1_3visit_draws - co2_mod1_draws) +
  scale_y_discrete(labels = mod1_3visit_plot_labels)

ch4_mod1_draws <- ch4_mod1$draws(variables = c("mu", "alpha_hyper", "alpha", "beta[1]", "beta[3]", "beta[5]", "gamma_hyper",
                                               "gamma[1]", "gamma[3]", "gamma[5]", "gamma[6]", "gamma[8]", "gamma[10]"))
ch4_mod1_3visit_draws <- ch4_mod1_3visit$draws(variables = c("mu", "alpha_hyper", "alpha", "beta", "gamma_hyper", "gamma"))
ch4_mod1_difference_plot <- bayesplot::mcmc_intervals(ch4_mod1_3visit_draws - ch4_mod1_draws)  +
  scale_y_discrete(labels = mod1_3visit_plot_labels) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

n2o_mod1_draws <- n2o_mod1$draws(variables = c("mu", "alpha_hyper", "alpha", "beta[1]", "beta[3]", "beta[5]", "gamma_hyper",
                                               "gamma[1]", "gamma[3]", "gamma[5]", "gamma[6]", "gamma[8]", "gamma[10]"))
n2o_mod1_3visit_draws <- n2o_mod1_3visit$draws(variables = c("mu", "alpha_hyper", "alpha", "beta", "gamma_hyper", "gamma"))
n2o_mod1_difference_plot <- bayesplot::mcmc_intervals(n2o_mod1_3visit_draws - n2o_mod1_draws) +
  scale_y_discrete(labels = mod1_3visit_plot_labels) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

ggarrange(co2_mod1_difference_plot, ch4_mod1_difference_plot + rremove("ylab"), n2o_mod1_difference_plot + rremove("ylab"),
          labels = c("CO2", "CH4", "N2O"),
          nrow = 1,
          widths = c(0.4,0.3,0.3))

```
We can see here that every posterior difference median falls almost exactly at 0. This indicates that, at least when considering how greenhouse gases change over the year, reducing sampling effort from 5 to 3 will not make a difference on inference.

And now for the second modelling question:

```{r mod2_difference, echo = FALSE}
co2_mod2_draws <- co2_mod2$draws(variables = c("mu", "alpha_hyper", "alpha", "beta", "gamma_hyper",
                                               "gamma"))
co2_mod2_3visit_draws <- co2_mod2_3visit$draws(variables = c("mu", "alpha_hyper", "alpha", "beta", "gamma_hyper", "gamma"))
co2_mod2_difference_plot <- bayesplot::mcmc_intervals(co2_mod2_3visit_draws - co2_mod2_draws) +
  scale_y_discrete(labels = mod2_3visit_plot_labels)

ch4_mod2_draws <- ch4_mod2$draws(variables = c("mu", "alpha_hyper", "alpha", "beta", "gamma_hyper",
                                               "gamma"))
ch4_mod2_3visit_draws <- ch4_mod2_3visit$draws(variables = c("mu", "alpha_hyper", "alpha", "beta", "gamma_hyper", "gamma"))
ch4_mod2_difference_plot <- bayesplot::mcmc_intervals(ch4_mod2_3visit_draws - ch4_mod2_draws) +
  scale_y_discrete(labels = mod2_3visit_plot_labels) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

n2o_mod2_draws <- n2o_mod2$draws(variables = c("mu", "alpha_hyper", "alpha", "beta", "gamma_hyper",
                                               "gamma"))
n2o_mod2_3visit_draws <- n2o_mod2_3visit$draws(variables = c("mu", "alpha_hyper", "alpha", "beta", "gamma_hyper", "gamma"))
n2o_mod2_difference_plot <- bayesplot::mcmc_intervals(n2o_mod2_3visit_draws - n2o_mod2_draws) +
  scale_y_discrete(labels = mod2_3visit_plot_labels) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

ggarrange(co2_mod2_difference_plot, ch4_mod2_difference_plot + rremove("ylab"), n2o_mod2_difference_plot + rremove("ylab"),
          labels = c("CO2", "CH4", "N2O"),
          nrow = 1,
          widths = c(0.4,0.3,0.3))

```

For this particular question, we *do* see some differences from 0, but they are hardly enough to make a huge difference. All of the 90% credible intervals overlap 0, indicating very little evidence of a difference. Of note, one of the 50% credibles (recall that these are the bold lines) appears to not quite overlap 0; this is the ON12 site effect for N2O. Let's take a closer look:


```{r mod3_difference, echo = FALSE}
on12_draws_full <- n2o_mod2$draws("beta[12]")
on12_draws_reduced <- n2o_mod2_3visit$draws("beta[12]")

ggarrange(bayesplot::mcmc_intervals(on12_draws_full)+
  scale_y_discrete(labels = "Full"),
bayesplot::mcmc_intervals(on12_draws_reduced)+
  scale_y_discrete(labels = "Reduced"),
bayesplot::mcmc_intervals(on12_draws_reduced - on12_draws_full)+
  scale_y_discrete(labels = "Difference"),
ncol = 1)

```

So, it appears that for site ON12, there is some information missing from changing from 5 visits to 3 visits, for this particular data set. One option to explore further is to run a simulation study to simulate multiple different data collections to see if this is a difference that is often captured, or if this is just a chance data collection. For time sensitivity, I will not run one here, but will proceed with this particular site in mind.

# Recommendations

Based on the analyses above, I would be comfortable recommending reducing sampling efforts from 5 visits per site to 3 visits per site in 2025. This is because for most statistical effects that we might be interested in terms of site effects, year effects, and intra-annual effects, there was little to no difference in posterior distributions when considering the full data set compared to the reduced data set. This implies that either the three visits per site is informative enough, or our prior distributions are strong enough on our model to compensate for lower sample size.

One could further investigate modelling future data by generating simulated data sets of new data, and modelling this with the data already collected. However, this exercise would only be useful in a situation where there are specific effect sizes or posterior distributions that we are wanting to find. This would be akin to a power analysis in a frequentist approach, but with the caveat that a "power analysis" does not really exist in a Bayesian context.

One concern would be the sampling of N2O at site ON12. At least with this particular data set, there did seem to be a loss of information in the reduced data set, such that the estimated posterior distribution for the site-level effects were more different than any other site-level effects compared to the full data set. However, if the new data (i.e., 2025) and the old data (i.e., 2023 and 2024) were to be modelled together, **particularly in a Bayesian framework such as what was done here**, there should be enough information sharing within the model that any effects missed by skipping Visits 2 and 4 in 2025 may be able to be roughly estimated given information from Visits 2 and 4 in 2023 and 2024. Therefore, I would recommend that if it is decided to reduce sampling to 3 visits in 2025, that careful modelling of the data later on should be considered.